Rare-event sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body

### 2019-10-19 draft_p2.tex
Add Sarah's comments for intro and proofread whole paper

\begin{abstract}
This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently. 
%
Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
This work is the first application to fluid-structure interaction problems. 
%
The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times. 
%
Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme drag events of a turbulent flow on a bluff body. 
\end{abstract}

@@

% general comments on physical problem %
%
Turbulent flows are important in a variety of natural phenomena, industrial and civil applications with the characteristic feature to spontaneously develop intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}.
``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts.
Such anomalous constraints can threaten the structural integrity of embedded structures.
%

From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is not successful.
Therefore, simulation offers a practical approach to gain physical insight into these events, quantify their intensity and estimate their frequency of occurrence.
However, this requires very long simulations since these events are also very rare.
% general def on rare-event sampling %
%
{Rare-event sampling refer to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
%
In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is conducting by using both very long time-series (direct sampling) and rare-event sampling techniques.
%

% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
%
In fluid turbulence, rare-event sampling has been approached mainly from simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}. In that case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
%
An alternative approach is based on \emph{instantons} \cite{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise). Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
An example is the investigation of the physics of rogue waves~\cite{dematteis2018rogue,dematteis2019experimental}.
%
A drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
%

%
In this paper, a more general approach is considered for complex, possibly deterministic dynamical systems.
It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and designed to sample rare events of some observable with a higher frequency.
%    
%
Even though such ideas date back to the early 1950's, they have received ever-growing interest during the last 20 years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} or communication networks simulation \cite{villen-altamirano_restart:_1994}.
More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\cite{wouters2016rare}.
%
%
An original contribution of the present work is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom. 
%Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
% 
Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
%

The \acl{ams} algorithm \cite{cerou_adaptive_2007} inherits from old ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.  
In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}. The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016}. More recently is has been applied to rare events in stochastic models of wall-turbulence \cite{rolland_extremely_2018} and atmospheric dynamics \cite{bouchet2019rare}.

During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \cite{touchette_large_2009}.
Alongside, numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
Among them, the \ac{gktl} algorithm \cite{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \cite{giardina_simulating_2011,Laffargue_2013}.
%
{Recently, \ac{gktl} has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. Along the same line, rare-event sampling algorithms are here pushed aside traditional applications by considering fluid-structure interaction in a turbulent flow.}
%

{The paper is organized in two parts. The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow. This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.}
%
The motivation is twofold.
Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations. This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible gain obtained from them. This assessment is developed in the second part of the paper.

% Annonce du plan
In next section~\ref{sec:test_flow}, the flow set-up is introduced and the dynamics related to typical
drag fluctuations is described.
The statistical properties of the drag are then discussed.
In section~\ref{sec:direct_sampling}, the phenomenology of extreme fluctuations of extreme
drag fluctuations is investigated based on a direct sampling approach.
Both the instantaneous drag and time-averaged drag are considered.
It is found that sampled extreme events for the instantaneous drag share very similar dynamics. Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.
Next, section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforward. This difficulty is put in perspective with the phenomenology developed in the previous sections.
Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
This latter allows an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
As a specific application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
Perspectives and conclusion finalize this work.

@@

% introduce the flow 
%
The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
% why this flow?
%
Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost, and allow for a systematic study. 
% 
We believe that this system embeds the characteristic features that makes the application of rare-event algorithms both relevant and challenging for fluid-structure-interaction problems. 
%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
%
% rapid description of the flow
Turbulent eddies generated in the near-wake of the grid are carried downstream.
They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
%
%Finally, the obstacle does not deform or move either. 
%
%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
%

% lattice Boltzmann method 
%
The flow dynamics is integrated by the lattice Boltzmann (LB) method in our numerical simulations. 
While traditional methods in \emph{computational fluid dynamics} rely on a discretization of the Navier-Stokes equations, the LB method considers the fluid at a mesoscopic level. Capturing the dynamics of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs. 
%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level. Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
Further details about the LB method are given in Appendix and references therein.
In our context, this numerical method has been chosen principally for its computational manageability and efficiency.

% geometry
%
The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points. The squared obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel. The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}). 
% boundary conditions
%
No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure \cite{lbm_book}.  
%
Upstream the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as inlet condition. 
The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively. The initial distributions are imposed at equilibrium (see Appendix). 
In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$. The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the LB method. 
Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced. Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density. The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4\tau_0$; $\tau_c$ will be identified later as the correlation time of the drag signal.
	The vorticity is given in lattice units.}
	\label{fig:typical_vorticity}
\end{figure}

@@

% drag signal
%
The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
\begin{equation}
\label{eq:drag_definition}
f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
\end{equation}
where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix). 
Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
%
Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
As a consequence, the drag can eventually be expressed as the difference 
\begin{equation}
\label{eq:drag_approx}
f_d(t) = p_{fb}(t) - p_{base}(t)
\end{equation}
between the pressure integrated over the forebody, $p_{fb}(t)$, and the base of the obstacle, $p_{base}(t)$.
Pressure fluctuations are related to the dynamics of the vorticity field with regions of strong vorticity corresponding to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
%
% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
%

% Definition of turnover time
%
The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
\begin{equation}
\label{eq:turnover_time}
\tau_0 = \frac{R}{U},
\end{equation}
where $R$ is the size of the square and $U$ is the averaged velocity. 
%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
%
% typical fluctuation of the drag
%
Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field
in the vicinity of the base is only slightly perturbed.

@@

% drag signal
%
% Description du signal de trainee typique
Fig.~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square, $f_d(t)$, over five hundred turnover times.
The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
Therefore, it is  natural to model the drag as a (scalar) random process.

@@

% drag statistics
% correlation time
Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}. It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$. One can then argue that the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
%
This observation is important for the application of rare-event algorithms as it  will be discussed in section~\ref{sec:rare_events_algorithms}.
%
In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag. The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}. The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers.

@@

How rare is a fluctuation $a$ is quantified by its return time $r(a)$. We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
%
%
%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
one might expect that 
\begin{equation}
 \label{eq:return_time}
r(a) \underset{a\to\infty}{\propto} \exp(-la)
\end{equation}
where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$ \cite{lestang_computing_2018}. It is consistently found that the return time $r(a)$ is well approximated by an exponential for large levels $a$. Let us also point out some deviation from the exponential law at the largest levels, which are probably the consequence of undersampling.

@@

% intro
%
We have extracted  the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$. This set will be considered as representative of \emph{extreme events} in the upcoming study. The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and possibly identify generic features.
%
According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation. Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified. Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
%
In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.

>>>>

\begin{abstract}
This study evaluates the relevance of rare-event sampling techniques to accelerate the simulation of extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body.
The main idea is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned in order to sample large-amplitude events more frequently. 
%
Such techniques have been shown to be efficient for a wide range of problems in statistical physics, computer science, biochemistry, enabling the simulation of rare  events otherwise out of reach by direct sampling.
This work is the first application to fluid-structure interaction problems. 
%
The drag experienced by a squared obstacle placed in a turbulent flow (in two dimensions) is taken as a representative case study to investigate the performance of two major rare-event sampling algorithms, namely the \ac{ams} and the \ac{gktl} algorithms.
Practical evidence is given that the fast sweeping-time of fluid structures past the obstacle has a drastic influence on the efficiency of these two algorithms.
While it is shown that the \ac{ams} algorithm does not yield significant run-time savings, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of the time-averaged drag and estimate related statistics such as return times. 
%
Beyond the study of applicability of rare-event sampling techniques to a fluid-mechanical problem, this work also includes a detailed phenomenological description of extreme-drag events of a turbulent flow on a bluff body. 
\end{abstract}

@@

% general comments on physical problem %
%
Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \cite{lesieur_book,donzis_sreenivasan_2010,Yeung}.
``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts.
Such anomalous constraints can threaten the structural integrity of embedded structures.
%

From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \cite{KRAICHNAN}.
The use of perturbative methods in identifying resonant interactions among degrees of freedom responsible for extreme fluctuations is unsuccessful.
Therefore, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
However, this requires very long simulations since these events are rare.
% general def on rare-event sampling %
%
{Rare-event sampling refer to a large body of methods that aim at exploring preferentially regions of phase space corresponding to rare events, that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
%
In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is conducting by using both very long time-series (direct sampling) and rare-event sampling techniques.
%

% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
%
In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \cite{bec_burgers_2007}. In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \cite{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
%
An alternative approach is based on instantons \cite{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise). Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \cite{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
An example is the investigation of the physics of rogue waves~\cite{dematteis2018rogue,dematteis2019experimental}.
%
A drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
%

%
In this paper, a more general approach is considered for complex, possibly deterministic, dynamical systems.
It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories, and is designed to sample rare events of some observable with a higher frequency.
%    
%
Even though such ideas date back to the early 1950s, they have received ever-growing interest over the the last twenty years with successful applications in various domains such as chemistry \cite{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \cite{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \cite{louvin2017}, nonlinear dynamical systems \cite{tailleur_probing_2007} and communication networks simulation \cite{villen-altamirano_restart:_1994}.
More importantly, these type of algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\cite{wouters2016rare}.
%
%
An original contribution of the present work is certainly the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom. 
%Let us remark that most of previous successful applications involve dynamical systems with low-dimensional attractors. \EL{is it TRUE? \#11}
% 
Two different algorithms suitable for out-of-equilibrium dynamics are considered and compared. Namely, the Adaptive Multilevel Splitting algorithm and the Giardina-Kurchan-Tailleur-Lecomte algorithm.
%

The \acl{ams} algorithm \cite{cerou_adaptive_2007} builds on previous ideas about splitting approaches \cite{KahnHarris1951,glasserman_look_1998,glasserman_multilevel_1999} -- a detailed description will be given later.  
In recent years, it has allowed for the computation of rare events in problems involving a large number of degrees of freedom such as molecular dynamics simulations \cite{aristoff_adaptive_2015,teo_adaptive_2016}. The first use of the \ac{ams} algorithms for more complex dynamics, for instance stochastic partial differential equations was for the computation of rare trajectories in the Allen-Cahn equations \cite{rolland_computing_2016}. More recently is has been applied to rare events in stochastic models of wall-turbulence \cite{rolland_extremely_2018} and atmospheric dynamics \cite{bouchet2019rare}.

During the last decade, the main theoretical framework for the study of rare events in statistical physics has been the theory of large deviations \cite{touchette_large_2009}.
Alongside numerical methods have been developed to sample rare events \cite{moral_feynman-kac_2004}. 
Among them the \ac{gktl} algorithm \cite{giardina_direct_2006} is particularly suited for (chaotic) dynamical systems \cite{giardina_simulating_2011,Laffargue_2013}.
%
{Recently the \ac{gktl} algorithm has allowed to successfully perform the numerical simulation of extreme heat waves in a simplified modelling of the atmosphere \cite{ragone_computation_2018}. 
This achievement represents a significant leap in the applicability of rare-event sampling to complex dynamical systems. Along the same line, rare-event sampling algorithms are here pushed aside traditional applications to consider fluid-structure interaction in a turbulent flow.}
%

{The paper is organized in two parts. The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow. This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.}
%
Motivation for this study is twofold.
Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations. This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible gain obtained from them. This assessment is developed in the second part of the paper.

% Annonce du plan
In section~\ref{sec:test_flow}, the flow set-up is introduced and the dynamics related to typical
drag fluctuations is described.
The statistical properties of the drag are then discussed.
In section~\ref{sec:direct_sampling}, the phenomenology of extreme fluctuations of extreme
drag fluctuations is investigated based on a direct sampling approach.
Both the instantaneous drag and time-averaged drag are considered.
It is found that sampled extreme events for the instantaneous drag share very similar dynamics. Furthermore, extreme fluctuations for the time-averaged drag can be connected to the statistics of the instantaneous drag.
This feature is well supported by theoretical arguments applied to simplified stochastic dynamics.
Section~\ref{sec:rare_events_algorithms} reports the applicability of both the \ac{ams} and \ac{gktl} algorithms to the numerical simulation of extreme drag fluctuations, by using the same flow configuration.
In section~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforward. This difficulty is put in perspective with the phenomenology developed in the previous sections.
Section~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag, using the \ac{gktl} algorithm.
This latter allows an exceptional reduction of the computational cost required to simulate trajectories corresponding to extreme time-averaged drag values.
As a specific application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
Perspectives and conclusion end this work.

@@

% introduce the flow 
%
The drag exerted by a grid-generated turbulent flow onto a fixed squared obstacle is considered as a representative case study (see Fig.~\ref{fig:illustr_ecoulement}). 
% why this flow?
%
Although real-world applications would eventually imply three-dimensional dynamics, a simplified two-dimensional setting has been chosen here to reduce the computational cost and allow for a systematic study. 
% 
We believe that this system embeds the characteristic features that makes the application of rare-event algorithms both relevant and challenging for fluid-structure-interaction problems. 
%Namely, spatio-temporal chaos and the emergence of large-amplitude mechanical efforts on the structure.
%
% rapid description of the flow
Turbulent eddies generated in the near-wake of the grid are carried downstream.
They interact with each other and grow in size as expected for two-dimensional turbulent dynamics. 
The dimension of the grid is such that the size of the eddies that hit the square is comparable to its size, resulting in strong fluctuations of the drag acting on the square. 
%
%Finally, the obstacle does not deform or move either. 
%
%Through this simplified setting, our motivation is primarily to evaluate the operability of sampling techniques to capture extreme events with a significant run-time savings.  
%

% lattice Boltzmann method 
%
The flow dynamics is integrated by the \ac{lbm} in our numerical simulations. 
While traditional methods in computational fluid dynamics rely on a discretization of the Navier-Stokes equations, the \ac{lbm} considers the fluid at a mesoscopic level.
Capturing the dynamics of collections of fluid particles distributed on a lattice is here preferred to solving non-linear PDEs. 
%This seems crazy, however, most details at the mesoscopic level play actually no role at the macroscopic level. Therefore, the LB algorithm may be viewed as a minimal kinetic scheme compliant to the fluid dynamics at the macroscopic level.
Further details about the \ac{lbm} are given in Appendix and references therein.
In our context, this numerical method has been chosen principally for its computational manageability and efficiency.

% geometry
%
The simulated flow develops in a long plane channel of dimension $513 \times 129$ mesh points. The square obstacle has size $R=16$ (in mesh unit) and is located at the centre of the channel. The spacing and bar height of the entrance grid are both equal to $R/2$ (see Fig.~\ref{fig:illustr_ecoulement}). 
% boundary conditions
%
No-slip boundary conditions are enforced on top and bottom walls of the channel and on the surface of the obstacle by using an halfway bounce-back procedure~\cite{lbm_book}.  
%
Upstream of the grid, a constant parabolic velocity profile and a constant mass density (equal to unity) are imposed as an inlet condition. 
The centerline velocity is $0.05$ in lattice units, \textit{i.e.} normalised by $\Delta x$ and $\Delta t$ referring to the lattice resolution and the time-step respectively. The initial distributions are imposed at equilibrium (see Appendix). 
In the bulk, the viscosity is adjusted so that grid turbulence is generated with Reynolds number $\mathrm{Re_{grid}}=1200$. The reference Mach number is equal to $0.06$ in agreement with the assumption of weak compressibility of the \ac{lbm}. 
Near the end of the channel, the flow is progressively damped within a \textit{sponge layer} where the viscosity is artificially enhanced.
Finally, the outlet boundary condition relies on a second-order extrapolation of the velocity and mass density.
The extrapolated distributions are evaluated through a regularization procedure relying on a finite difference estimation of the local stress tensor, as introduced in \cite{latt2008straight}.

@@

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{ecoulement_typique/ecoulement_typique.png}
	\caption{Snapshots of the vorticity related to typical drag fluctuations (within one standard deviation) over a time interval of length $\tau_c \simeq 4\tau_0$; $\tau_c$ will later be identified as the correlation time of the drag signal.
	The vorticity is given in lattice units.}
	\label{fig:typical_vorticity}
\end{figure}

@@

% drag signal
%
The incoming turbulent flow exerts fluctuating mechanical efforts onto the squared obstacle.
The \textit{drag} is defined as the resulting force in the streamwise $x$-direction. Formally 
\begin{equation}
\label{eq:drag_definition}
f_d(t) = \int_{\mathcal{S}} \boldsymbol{\tau}_{x \beta}(\mathbf{x},t) ~ \mathrm{d}{\mathcal{S}}_\beta(\mathbf{x}),
\end{equation}
where $\mathcal{S}$ is the surface of the obstacle and $\boldsymbol{\tau}$ denotes the stress tensor (see Appendix). 
Here, the viscous stress makes a negligible contribution to the drag. The latter therefore results mostly from pressure forces.
% which are closely related to the distribution of velocity gradients in the vicinity of the obstacle (in the nearly-incompressible limit). 
%
Since the pressure on the top and bottom sides of the square applies in the normal direction, they do not contribute to the drag. 
As a consequence, the drag can eventually be expressed as the difference 
\begin{equation}
\label{eq:drag_approx}
f_d(t) = p_{fb}(t) - p_{base}(t)
\end{equation}
between the pressure integrated over the upstream side of the obstacle or \textit{forebody}, $p_{fb}(t)$, and the downstream side or \textit{base} $p_{base}(t)$.
Pressure fluctuations are related to the dynamics of the vorticity fiel.
Regions of strong vorticity correspond to strong local pressure gradients, \emph{e.g.} as demonstrated analytically with a Rankine vortex.
%
% Considering a disc of constant vorticity $\omega$ with radius $a$ (Rankine vortex model), the local pressure gradient between the vortex core and a point infinitely far grows quadratically with the vorticity.
%

% Definition of turnover time
%
The typical timescale (turnover time) of drag fluctuations can be estimated from dimensional analysis as
\begin{equation}
\label{eq:turnover_time}
\tau_0 = \frac{R}{U},
\end{equation}
where $R$ is the size of the square and $U$ is the averaged velocity in the channel. 
%The viscosity is here discarded since viscous stress is negligible as compared to pressure forces.
%
% typical fluctuation of the drag
%
Fig.~\ref{fig:typical_vorticity} displays the typical evolution of the vorticity field around the obstacle over a few turnover times.
Because the vorticity generated along the forebody is swept away by the mean flow, the pressure field in the vicinity of the base is only slightly perturbed.

@@

% drag signal
%
% Description du signal de trainee typique
Figure~\ref{fig:typical_drag_signal} shows the time signal of the drag acting on the square, $f_d(t)$, over five hundred turnover times.
The signal appears unpredictable in details and exhibits repeated bursts of high amplitude that deviate significantly from the averaged value.
Therefore, it is natural to model the drag as a (scalar) random process.

@@

% drag statistics
% correlation time
Lastly, the autocorrelation function of the drag $C(\tau)$ is shown in Fig.~\ref{fig:pdf_drag_b}. It is found that drag fluctuations are correlated over a time interval $\tau_c \simeq 4\tau_0$. One can then argue that the drag looses its memory  over a time scale corresponding to the sweeping of a few eddies past the obstacle.
%
This observation is important for the application of rare-event algorithms as it will be discussed in section~\ref{sec:rare_events_algorithms}.
%
In the following, $\tau_c$ will be referred to as the \textit{correlation time} of the drag process.
The ratio $\tau_0 / \tau_c$ may be viewed as a {Strouhal number}. The value $St=0.25$ is consistent with common observations for flows past blunt structures at comparable Reynolds numbers.

@@

How rare is a fluctuation $a$ is quantified by its return time $r(a)$.
We can define extreme drag fluctuations as \textit{rare events} in the sense that the return time is much larger than the correlation time, \emph{i.e.} $r(a) \gg \tau_c$.
%
%
%The shape of the return time $r$ as a function of the return level $a$ is system dependent. 
If one assumes that  $r(a) = t(a)~/~\mathbb{P}(f_d\geq a)$   where the time scale $t(a)$ is of order $\tau_c$ and varies much more slowly with $a$ than ${\mathbb{P}(f_d\geq a)}$,
% As the PDF ${\mathbb{P}(f_d\geq a)}$ was shown to be well approximated by an exponential (see figure Fig.~\ref{fig:pdf_drag})
one might expect that 
\begin{equation}
 \label{eq:return_time}
r(a) \underset{a\to\infty}{\propto} \exp(-la)
\end{equation}
where $l$ is the rate describing the positive tail of the \ac{pdf} of the drag (shown in Fig.~\ref{fig:pdf_drag}).
Fig.~\ref{fig:return_time_instant} shows the evolution of the return time $r(a)$ with the amplitude of fluctuation $a$, computed from {direct sampling} of the drag signal $f_d(t)$~\cite{lestang_computing_2018}.
Consistently, it is found that the return time $r(a)$ is well approximated by an exponential for large levels $a$. Let us also point out some deviation from the exponential law at the largest levels, which are probably the consequence of under-sampling.

@@

% intro
%
We have extracted the fluctuations of the drag with a return time $r(a)$ greater than  $10^4\tau_c$ from the control time-series $\{f_d(t)\}_{0 \leq t \leq T_{tot}}$.
This set will be considered as representative of \emph{extreme events} in the upcoming study.
The choice of this particular threshold has been driven by the need to collect enough events with large amplitude and possibly identify generic features.
%
According to Fig.~\ref{fig:return_time_instant}, the related amplitude $a$ is found equal to $7.6~\sigma$ with $\sigma$ being the standard deviation of the drag process.
Precisely, 104 independent fluctuations with $f_d(t) \geq 7.6\sigma$ have been identified. Each fluctuation is characterized by its maximal value, $f_d^{\star}$, and the time, $t^{\star}$, at which this maximum is reached.
%
In the following, the phenomenology of extreme drag fluctuations will be examined on the basis of this set of events.
